### 2009-03-02 doc/storage_system/arc-storage-documentation.tex
updating the storage documentation

\begin{verbatim}
    <Service name="pythonservice" id="librarian">
        <ClassName>storage.librarian.librarian.LibrarianService</ClassName>
        <AHashURL>https://localhost:60000/AHash</AHashURL>
        <HeartbeatTimeout>30</HeartbeatTimeout>
        <CheckPeriod>20</CheckPeriod>
        <ClientSSLConfig FromFile="clientsslconfig.xml"/>
        <TrustManager FromFile='librarian_trust_manager.xml'/>
    </Service>
\end{verbatim}

@@

@@

@@

The Bartender provides an easy to use interface of the ARC storage system to the users. You can put, get and delete files using their logical names (\emph{LN}s) with the \emph{putFile}, \emph{getFile} and \emph{delFile} methods, create, remove and list collections with \emph{makeCollection}, \emph{unmakeCollection} and \emph{list}. The metadata of a file or collection (e.g.~whether the collection is closed, number of needed replicas, access policies) can be changed with \emph{modify}. A \emph{stat} call gives all the information about a file or collection, and you can move (or hardlink) collections and files within the namespace with \emph{move}. You can upload an entirely new replica to a file (e.g.~if the file lost all its replicas, or when a Shepherd service offers its replica for replications) with \emph{addReplica}.

@@

\subsection{Prototype status and plans} % (fold)

The methods mentioned in the above section are all implemented, but need more error-checking and metadata-checking. There are plans of adding new methods, e.g.~a \emph{copy} method or a \emph{glob} method for file pattern matching. The current version does not support closed (unmodifiable) collections yet.

% subsection protoype_status_and_plans (end)

\subsection{Data model} % (fold)

The Bartender interface uses mostly Logical Names (\emph{LN}s), which have the syntax of: \verb!<GUID>/<path>! where both sides can be omitted (and in the case of a sole GUID we don't need the slash either), e.g.~\verb!afg342/foo! is an entry called \verb!foo! in the collection with GUID \verb!afg342!; the LN \verb!f36a7481! refers to the a file or collection with GUID \verb!f36a7481!; \verb!/vo/dir/stg! points to the entry which is reachable from the root collection using the given path; and \verb!/! simply refers to the root collection.
The term `\emph{metadata}' here refers to a list of property-value pairs organized in sections, see the data model description in Section~\ref{sec:data_model}.

% subsection data_model (end)

@@

\begin{description}
    \item[putFile(putFileRequestList)] returns a list of (\emph{requestID}, \emph{success}, \emph{TURL}, \emph{protocol})

    \emph{putFileRequestList} is a list of (\emph{requestID}, \emph{LN}, \emph{metadata}, \emph{protocols}), where \emph{requestID} is an arbitrary ID which will be used in the response; \emph{LN} is the chosen Logical Name of the new file, \emph{protocols} is a list of protocols we can use for uploading, \emph{metadata} is a list of (\emph{section}, \emph{property}, \emph{value}) tuples where properties could be in the `\textbf{states}' section: `\textbf{size}', `\textbf{checksum}', `\textbf{checksumType}', and  `\textbf{neededReplicas}', policy documents in the `\textbf{policies}' section and any other property-value pairs in the `\textbf{metadata}' section. The returned \emph{TURL} is a URL with a chosen \emph{protocol} to upload the file itself, the \emph{success} string could be `\textbf{done}', `\textbf{missing metadata}', `\textbf{parent does not exists}', `\textbf{internal error:} reason', etc.

    \item[getFile(getFileRequestList)] returns a list of (\emph{requestID}, \emph{success}, \emph{TURL}, \emph{protocol})

    \emph{getFileRequestList} is a list of (\emph{requestID}, \emph{LN}, \emph{protocols}) where \emph{requestID} is used in the response, \emph{LN} is the Logical Name referring to the file we want to get, \emph{protocols} is a list of transfer protocols the client supports.
    In the response \emph{TURL} is the transfer URL using \emph{protocol}, with which we can download the file, \emph{success} could be `\textbf{done}', `\textbf{not found}', `\textbf{is not a file}', `\textbf{file has no valid replica}', `\textbf{error while getting TURL:} reason', etc.

    \item[delFile(delFileRequestList)] returns a list of (\emph{requestID}, \emph{status})

    \emph{delFileRequestList} is a list of (\emph{requestID}, \emph{LN}) with the Logical Name of the file we want to delete. The status in response could be `\textbf{deleted}' or `\textbf{nosuchLN}'.

    \item[stat(statRequestList)] returns a list of (\emph{requestID}, \emph{metadata})

    \emph{statRequestList} is a list of (\emph{requestID}, \emph{LN}) with the Logical Name of the file or collection we want to get information about, and it returns \emph{metadata} which is a list of (\emph{section}, \emph{property}, \emph{value}) tuples according to the data model of the Librarian (see Section~\ref{sec:data_model})

    \item[makeCollection(makeCollectionRequestList)] returns a list of (\emph{requestID}, \emph{success})

    \emph{makeCollectionRequestList} is a list of (\emph{requestID}, \emph{LN}, \emph{metadata}) where \emph{metadata} is a list of (\emph{section}, \emph{property}, \emph{value}) tuples where in the `\textbf{entries}' section there could be the initial content of the catalog in the form of name-GUID pairs (these entries will be hard links to the given GUIDs with the given name), in the `\textbf{states}' section there is the `\textbf{closed}' property (if it is true then no more files can be added or removed later), in the `\textbf{policies}' section there could be some access policies, and in the `\textbf{metadata}' section there could be any other metadata in property-value pairs. The \emph{success} in the response could be `\textbf{done}', `\textbf{LN exists}', `\textbf{parent does not exist}', `\textbf{failed to create new catalog entry}', `\textbf{failed to add child to parent}', `\textbf{internal error}', etc.
    \item[unmakeCollection(unmakeCollectionRequestList)] returns a list of (\emph{requestID}, \emph{success}).

    \emph{unmakeCollectionRequestList} is a list of (\emph{requestID}, \emph{LN}) with the Logical Names of the collections we want to remove. \emph{success} could be `\textbf{removed}', `\textbf{no such LN}', `\textbf{collection is not empty}', `\textbf{failed}: reason'.

    \item[list(listRequestList, neededMetadata)] returns \emph{listResponse}.

    \emph{listRequestList} is a list of (\emph{requestID}, \emph{LN}) where \emph{LN} is the Logical Name of the collection (or file) we want to list, \emph{neededMetadata} is a list of (\emph{section}, \emph{property}) pairs which filters the returned metadata.

    \emph{listResponse} is a list of (\emph{requestID}, \emph{entries}, \emph{status}) where entries is a list of (\emph{name}, \emph{GUID}, \emph{metadata}) where \emph{metadata} is a list of (\emph{section}, \emph{property}, \emph{value}) tuples according to the data model of the Librarian (Section~\ref{sec:data_model}), the \emph{status} could be `\textbf{found}', `\textbf{not found}', `\textbf{is a file}' (because only collections can be listed).

    \item[move(moveRequestList)] returns a list of (\emph{requestID}, \emph{status}).

    \emph{moveRequestList} is a list of (\emph{requestID}, \emph{sourceLN}, \emph{targetLN}, \emph{preserveOriginal}) where \emph{sourceLN} is the Logical Name referring to the file or collection we want to move (or just rename) and \emph{targetLN} is the new path, and if \emph{preserveOriginal} is true the \emph{sourceLN} would not be removed, so with \emph{preserveOriginal} we actually creating a hard link. The status could be `\textbf{moved}', `\textbf{nosuchLN}', `\textbf{targetexists}', `\textbf{invalidtarget}', `\textbf{failed adding child to parent}', `\textbf{failed removing child from parent}'

    \item[modify(modifyRequestList)] returns a list of (\emph{changeID}, \emph{success})

    \emph{modifyRequestList} is a list of (\emph{changeID}, \emph{LN}, \emph{changeType}, \emph{section}, \emph{property}, \emph{value}) where \emph{changeType} can be `\textbf{set}' (set the \emph{property} in the \emph{section} to \emph{value}), `\textbf{unset}' (remove the \emph{property}-\emph{value} pair from the \emph{section}), `\textbf{add}' (set the \emph{property} in the \emph{section} to \emph{value} only if it is not exists already). \emph{success} could be `\textbf{no such LN}', `\textbf{set}', `\textbf{unset}', `\textbf{entry exists}', `\textbf{failed}: reason'.

@@

>>>>

\begin{verbatim}
    <Service name="pythonservice" id="librarian">
        <ClassName>storage.librarian.librarian.LibrarianService</ClassName>
        <AHashURL>https://localhost:60000/AHash</AHashURL>
        <HeartbeatTimeout>30</HeartbeatTimeout>
        <CheckPeriod>20</CheckPeriod>
        <ClientSSLConfig FromFile="clientsslconfig.xml"/>
    </Service>
\end{verbatim}

@@

\subsection{Implementation} % (fold)
\label{sub:shepherd_implementation}

@@

\begin{description}
    \item[ServiceID] is the ID of the service, currently it should be the URL of the service, because there is no information system where we can discover services by IDs.
    \item[CheckPeriod] specifies in seconds how frequently should the Shepherd check the existence and the checksum of all the files on the storage node.
    \item[MinCheckInterval] specifies in seconds how much time should the Shepherd wait between two subsequent file checks. This ensures that the Shepherd does not use up all the resources of the hosting machine. After checking the cheksum of a file, the Shepherd will always wait this long before checking the next file, even if this means that checking all the files takes longer than the `CheckPeriod'.
    \item[StoreClass] specifies which type of store the Shepherd should use for storing its metadata on disk.
    \item[StoreCfg] can have parameters for the store class, e.g. the local directory to be used.
    \item[BackendClass] specifes which backend the Shepherd should use for communicating with the storage element service.
    \item[BackendCfg] has some configuration parameters for the backend class: \textbf{DataDir} and \textbf{TransferDir} is the directories the backend will use (see Section~\ref{sub:shepherd_implementation}), \textbf{TURLPrefix} should be the URL of the storage element service ending with a `/'.
    \item[LibrarianURL] is the URL of a Librarian where the Shepherd should send the reports.
    \item[BartenderULR] is the URL of a Bartender whom the Shepherd should connect if it wants to offer a replica for replication or recover a corrupted replica.
\end{description}

@@

The Bartender provides an easy to use interface of the ARC storage system to the users. You can put, get and delete files using their logical names (\emph{LN}s) with the \emph{putFile}, \emph{getFile} and \emph{delFile} methods, create, remove and list collections with \emph{makeCollection}, \emph{unmakeCollection} and \emph{list}. You can remove the link to a file or sub-collection from its parent collection without removing the file or sub-collection itself with the \emph{unlink} method. The metadata of a file or collection (e.g.~whether the collection is closed, number of needed replicas, access policies) can be changed with \emph{modify}. A \emph{stat} call gives all the information about a file or collection, and you can move (or create hardlinks of) collections and files within the namespace with \emph{move}. You can upload an entirely new replica to a file (e.g.~if the file lost all its replicas, or when a Shepherd service offers its replica for replications) with \emph{addReplica}. You can create and remove mount points with the \emph{makeMountpoint} and \emph{unmakeMountpoint} methods.

The Bartender communicates with the Librarians to get and modify the metadata of files and collections, and it communicates with the Shepherds to initiate file transfers. The Bartender also has \emph{gateway modules} which are capable of communicating different kinds of thrid-party storage solutions. With these gateway modules it is possible to create mount points within the namespace of the ARC storage, and to access the namespace of the third-party storage through these mount points, which means the user can use the Bartender to get listing of directories on the third-party storage or to get files from the third-party storage. Thus the user can use a single client tool to access different kind of storages easily.

The high-level authorization is done by the Bartender. It makes decisions based on the identity of the connecting client and the access policy rules of the files and collections. The Bartender uses the security framework of the HED to evaluate the requests and the policies, which are genereted by the Bartender from the internal representation of the access rules. For more details see Section~\ref{sec:security_implementation}.

@@

@@

    \item[putFile(putFileRequestList)] creates new files at the requested Logical Names, chooses a Shepherd and initiates the file transfer, and returns a TURL for each new file.

    The \emph{putFileRequestList} is a list of (\emph{requestID}, \emph{LN}, \emph{metadata}, \emph{protocols}), where \emph{requestID} is an arbitrary ID which will be used in the response; \emph{LN} is the requested Logical Name of the new file, \emph{protocols} is a list of protocols we can use for uploading, \emph{metadata} is a list of (\emph{section}, \emph{property}, \emph{value}) tuples which should contain some mandatory properties in the `\textbf{states}' section: `\textbf{size}', `\textbf{checksum}', `\textbf{checksumType}', and  `\textbf{neededReplicas}'.

    The response is a list of (\emph{requestID}, \emph{success}, \emph{TURL}, \emph{protocol}), where \emph{TURL} is a URL with a chosen \emph{protocol} which we can use to upload the file, the \emph{success} string could be `\textbf{done}', `\textbf{missing metadata}', `\textbf{parent does not exists}', `\textbf{internal error:} reason', etc.

    \item[getFile(getFileRequestList)] finds the files referenced by the given Logical Name, chooses a replica and initiates the download, and returns a TURL for each file.

    The \emph{getFileRequestList} is a list of (\emph{requestID}, \emph{LN}, \emph{protocols}) where \emph{requestID} is used in the response, \emph{LN} is the Logical Name referring to the file we want to get, \emph{protocols} is a list of transfer protocols we support.

    The response is a list of (\emph{requestID}, \emph{success}, \emph{TURL}, \emph{protocol}), where \emph{TURL} is the transfer URL using \emph{protocol}, with which we can download the file, \emph{success} could be `\textbf{done}', `\textbf{not found}', `\textbf{is not a file}', `\textbf{file has no valid replica}', `\textbf{error while getting TURL:} reason', etc.

    \item[delFile(delFileRequestList)] removes the files references by the given Logical Names.

    The \emph{delFileRequestList} is a list of (\emph{requestID}, \emph{LN}) with the Logical Name of the files we want to delete.

    The response is a list of (\emph{requestID}, \emph{status}), where the \emph{status} could be `\textbf{deleted}' or `\textbf{nosuchLN}'.

    \item[unlink(unlinkRequestList)] remove a link from a collection without deleting the file or sub-collection itself.

    The \emph{unlinkRequestList} is a list of (\emph{requestID}, \emph{LN}) with the Logical Name of the files and sub-collections we want to unlink.

    The response is a list of (\emph{requestID}, \emph{success}) where \emph{success} can be `\textbf{unset}', `\textbf{no such LN}', `\textbf{denied}' or `\textbf{nothing to unlink}'

    TODO: makeMountpoint, unmakeMountpoint

    \item[stat(statRequestList)] returns all metadata of files, collections or mount points

    The \emph{statRequestList} is a list of (\emph{requestID}, \emph{LN}) with the Logical Names we want to get information about.

    The response is a list of (\emph{requestID}, \emph{metadata}), where \emph{metadata} is a list of (\emph{section}, \emph{property}, \emph{value}) tuples (see the data model in Section~\ref{sec:data_model}).

    \item[makeCollection(makeCollectionRequestList)] creates new collections.

    The \emph{makeCollectionRequestList} is a list of (\emph{requestID}, \emph{LN}, \emph{metadata}) where \emph{metadata} is a list of (\emph{section}, \emph{property}, \emph{value}) tuples where in the `\textbf{entries}' section there could be the initial content of the catalog in the form of name-GUID pairs (these entries will be hard links to the given GUIDs with the given name) and in the `\textbf{states}' section there is the `\textbf{closed}' property (if it is `yes' then no more files can be added or removed later).

    The response is a list of (\emph{requestID}, \emph{success}), where \emph{success} can be `\textbf{done}', `\textbf{LN exists}', `\textbf{parent does not exist}', `\textbf{failed to create new catalog entry}', `\textbf{failed to add child to parent}', `\textbf{internal error}', etc.

    \item[unmakeCollection(unmakeCollectionRequestList)] deletes empty collections.

    The \emph{unmakeCollectionRequestList} is a list of (\emph{requestID}, \emph{LN}) with the Logical Names of the collections we want to remove.

    The response is a list of (\emph{requestID}, \emph{success}), where \emph{success} could be `\textbf{removed}', `\textbf{no such LN}', `\textbf{collection is not empty}', `\textbf{failed}: reason'.

    \item[list(listRequestList, neededMetadata)] returns the contents of the requested collections.

    The \emph{listRequestList} is a list of (\emph{requestID}, \emph{LN}) where \emph{LN} is the Logical Name of the collection we want to list, \emph{neededMetadata} is a list of (\emph{section}, \emph{property}) pairs which filters the returned metadata.

    The response is a a list of (\emph{requestID}, \emph{entries}, \emph{status}) where entries is a list of (\emph{name}, \emph{GUID}, \emph{metadata}) where \emph{metadata} is a list of (\emph{section}, \emph{property}, \emph{value}) tuples (see the data model in Section~\ref{sec:data_model}); the \emph{status} could be `\textbf{found}', `\textbf{not found}', `\textbf{is a file}'.

    \item[move(moveRequestList)] moves file or collections within the namespace which changes the Logical Name of the file or collection. This method only alters metadata, does not move real file data.

    The \emph{moveRequestList} is a list of (\emph{requestID}, \emph{sourceLN}, \emph{targetLN}, \emph{preserveOriginal}) where \emph{sourceLN} is the Logical Name referring to the file or collection we want to move (rename) and \emph{targetLN} is the new path, and if \emph{preserveOriginal} is true the \emph{sourceLN} would not be removed, so with \emph{preserveOriginal} we are actually creating a hard link.

    The response is a list of (\emph{requestID}, \emph{status}), where status could be `\textbf{moved}', `\textbf{nosuchLN}', `\textbf{targetexists}', `\textbf{invalidtarget}', `\textbf{failed adding child to parent}', `\textbf{failed removing child from parent}'.

    \item[modify(modifyRequestList)] modifies the metadata of files, collections and mount poins. Only the `states', the `policy' and the `metadata' sections can be modified, and there are seperate access control actions for the three.

    The \emph{modifyRequestList} is a list of (\emph{changeID}, \emph{LN}, \emph{changeType}, \emph{section}, \emph{property}, \emph{value}) where \emph{changeType} can be `\textbf{set}' (set the \emph{property} in the \emph{section} to \emph{value}), `\textbf{unset}' (remove the \emph{property}-\emph{value} pair from the \emph{section}), `\textbf{add}' (set the \emph{property} in the \emph{section} to \emph{value} only if it is not exists already).

    The response is a list of (\emph{changeID}, \emph{success}), where \emph{success} could be  `\textbf{set}', `\textbf{unset}', `\textbf{entry exists}' (for an `add' request), `\textbf{denied}', `\textbf{no such LN}', `\textbf{failed}: reason'.

@@

For managing collections the Bartender needs only to communicate with a Librarian service, for managing files it needs to communicate with Shepherd services as well. When a Bartender wants to create a new replica of a file (or the first replica of a new file) it should find available Shepherd services. Fortunately the Librarian has a list of registered Shepherd services, so the Bartender can use that.

TODO: The Bartender can utilize gateway modules for accessing file listings and transfer URL on third-party storages. Curretly we have one implementation of a gateway module [...] A gateway module should provide these methods: [...]

TODO: a few sentences about the proxy delegation.

The Bartender has these configuration variables:

\begin{description}
    \item[LibrarianURL] specifies the URL of one or more Librarian services.
    \item[ProxyStore] is a local directory where the Bartender will store the delegated proxies.
    \item[GatewayClass] is the name of the gateway module we want to use in this Bartender.
    \item[GatewayCfg] contains the configuration parameters of the gateway class: the \textbf{ProxyStore} is the directory where the gateway module could find delegated proxies, and the \textbf{CACertificatesDir} is a directory where the certificates of the trusted CAs are. 
\end{description}

\begin{verbatim}
    <Service name="pythonservice" id="bartender">
        <ClassName>storage.bartender.bartender.BartenderService</ClassName>
        <LibrarianURL>https://sal1.uppmax.uu.se:60000/Librarian</LibrarianURL>
        <ProxyStore>/var/arc/spool/proxy_store</ProxyStore>
        <GatewayClass>storage.bartender.gateway.gateway.Gateway</GatewayClass>
        <GatewayCfg>
            <ProxyStore>/var/arc/spool/proxy_store</ProxyStore>
            <CACertificatesDir>/etc/grid-security/certificates/
        </GatewayCfg>
        <ClientSSLConfig FromFile='clientsslconfig.xml'/>
    </Service>
\end{verbatim}
